---
title: "ceterisParibus2 plots for regression models"
author: "Dariusz Komosinski"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{ceterisParibus2 plots for regression models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  cache = TRUE
)
```

Here we will use the dragons data to present the ceterisParibus2 package for regression models.

```{r}
# devtools::install_github("ModelOriented/DALEX2")
library(DALEX2)
library(ggplot2)
library(ceterisParibus2)
  
head(dragons)
new_observation <- dragons_test[1,]
new_observation
```

Select neighbourhood sample, random sample, specific variables.

```{r}
similar_dragons <- select_neighbours(dragons_test, new_observation, n = 10)
similar_dragons

random_dragons <- select_sample(dragons_test, n = 10)
random_dragons

variable_splits <- calculate_variable_split(dragons, variables = c("scars", "weight"))
```


## Linear regression

First, we fit a model.

```{r}
m_lm <- lm(life_length ~ . , data = dragons)
```

To calculete individual variable profiles (ceteris paribus profiles), i.e. series of predictions from a model calculated for observations with altered single coordinate, we use the `ceterisParibus2` package.

We create an object of the `ceteris_paribus_explainer` class.

```{r}
ivp_lm <- individual_variable_profile(m_lm,
                            data = dragons_test,
                            new_observation =  new_observation)
```

Now we can plot individual profiles
```{r}
plot(ivp_lm)
```

Calculate oscillations for individual variable profiles.

```{r}
individual_variable_oscillations(ivp_lm)
```

Ceteris paribus plots for neighbourhood sample, random sample, selected variables.

```{r}
ivp_lm_neighbours <- individual_variable_profile(m_lm,
                            data = similar_dragons,
                            new_observation =  new_observation)
plot(ivp_lm_neighbours)

ivp_lm_random <- individual_variable_profile(m_lm,
                            data = random_dragons,
                            new_observation =  new_observation)
plot(ivp_lm_random)
```

Data frame with profiles.

```{r}
profiles <- calculate_variable_profile(data = new_observation,
                                       variable_splits = variable_splits,
                                       model = m_lm)
head(profiles)
```

For another type of models we proceed analogously.

## randomForest

```{r}
library(randomForest)

m_rf <- randomForest(life_length ~ . , data = dragons)

ivp_rf <- individual_variable_profile(m_rf,
                            data = dragons_test,
                            new_observation =  new_observation)

plot(ivp_rf)
individual_variable_oscillations(ivp_rf)
```

## SVM

```{r}
library(e1071)

m_svm <- svm(life_length ~ . , data = dragons)

ivp_svm <- individual_variable_profile(m_svm,
                            data = dragons_test,
                            new_observation =  new_observation)

plot(ivp_svm)
individual_variable_oscillations(ivp_svm)
```

## knn

```{r}
library(caret)

m_knn <- knnreg(life_length ~ . , data = dragons, k = 5)

ivp_knn <- individual_variable_profile(m_knn,
                            data = dragons_test,
                            new_observation =  new_observation)

plot(ivp_knn)
individual_variable_oscillations(ivp_knn)
```

## nnet

When you use `nnet` package for regression, remember to normalize the resposne variable, in such a way that it is from interval $(0,1)$.

In this case creating own predict function is also needed.

```{r}
library(nnet)

x <- max(abs(dragons$life_length))
digits <- floor(log10(x))
normalizing_factor <- round(x, -digits)
m_nnet <- nnet(life_length/normalizing_factor ~ . , data = dragons, size = 10, linout = TRUE)

p_fun <- function(model, new_observation){
  predict(model, newdata = new_observation)*normalizing_factor
}

ivp_nnet <- individual_variable_profile(m_nnet,
                            data = dragons_test,
                            new_observation =  new_observation,
                            predict_function = p_fun)

plot(ivp_nnet)
individual_variable_oscillations(ivp_nnet)
```

## Several models at once

To produce plot with many models in one graph, use argument `color = _label_`.

```{r}
plot(ivp_lm, ivp_rf, ivp_svm, ivp_knn, ivp_nnet, color = "_label_")
```


<!-- ## lightGBM -->

<!-- ```{r} -->


<!-- ``` -->
